---
title: "Thesis Results - PMA"
output:
  html_notebook:
    toc: yes
  html_document:
    toc: yes
    df_print: paged
---

# Load libraries
```{r}
# install.packages("ppcor")
# install.packages("apaTables")

library(ppcor)
library(psych)
library(tidyverse)
library(apaTables)
```


# Read in data
```{r}
PM <- read_csv("PM_Data_removed.csv")

```

# Describe data
```{r}
PM_Describe <- describe(PM)
PM_Describe
#apa.cor.table(PM, "PM_Describe.doc", show.conf.interval = TRUE)
```

# Hypothesis 1


```{r}
PM_full <- select(PM, Accuracy, RT, STAIS, STAIT, BAI, N300, PP)

cor(PM_full)

apa.cor.table(PM_full, "PM_full.doc", show.conf.interval = TRUE)

cor.test( formula = ~ N300 + PP,
          data = PM_full)

```




## Make a scatterplot: BAI & Accuracy
```{r}
ggplot(PM, aes(x = BAI, y = Accuracy)) +
  geom_point() +
    geom_smooth(method = "lm")
```
## Make a scatterplot: BAI & RT
```{r}
ggplot(PM, aes(x = BAI, y = RT)) +
  geom_point() +
    geom_smooth(method = "lm")
```
## Make a scatterplot: STAI-S & Accuracy
```{r}
ggplot(PM, aes(x = STAIS, y = Accuracy)) +
  geom_point() +
    geom_smooth(method = "lm")
```
## Make a scatterplot: STAI-S & RT
```{r}
ggplot(PM, aes(x = STAIS, y = RT)) +
  geom_point() +
    geom_smooth(method = "lm")
```

## Make a scatterplot: STAI-T & Accuracy
```{r}
ggplot(PM, aes(x = STAIT, y = Accuracy)) +
  geom_point() +
    geom_smooth(method = "lm")
```

## Make a scatterplot: STAI-T & RT
```{r}
ggplot(PM, aes(x = STAIT, y = RT)) +
  geom_point() +
    geom_smooth(method = "lm")
```


## select variables of interest
```{r}
PM_acc <- select(PM, Accuracy, STAIS, STAIT, BAI)
PM_RT <- select(PM, RT, STAIS, STAIT, BAI)
```


## Get correlation of variables
### Via cor() function
```{r}
cor(PM_acc)
cor(PM_RT)

cor.test( formula = ~ Accuracy + STAIS,
          data = PM_acc)
cor.test( formula = ~ Accuracy + STAIT,
          data = PM_acc)
cor.test( formula = ~ Accuracy + BAI,
          data = PM_acc)
cor.test( formula = ~ RT + STAIS,
          data = PM_RT)
cor.test( formula = ~ RT + STAIT,
          data = PM_RT)
cor.test( formula = ~ RT + BAI,
          data = PM_RT)


```

### Via apa.cor.table()
```{r}
apa.cor.table(PM_acc, "PM_acc.doc", show.conf.interval = TRUE)
apa.cor.table(PM_RT, "PM_RT.doc", show.conf.interval = TRUE)
```

## FDR correction
```{r}
# Vector of unadjusted p values 

p1 <- c(0.1252, 0.3723, 0.9245, 0.2691, 0.1776, 0.3924)

# Key:  Accuracy + STAIS, Accuracy + STAIT, Accuracy + BAI, RT + STAIS, RT + STAIT, RT + BAI

# FDR correction

FDR_Hyp1 <- p.adjust(p1, method = "fdr", n = length(p1))

FDR_Hyp1
```


# Hypothesis 2

## Make a scatterplot: N300 & BAI
```{r}
ggplot(PM, aes(x = N300, y = BAI)) +
  geom_point() +
    geom_smooth(method = "lm")
```

## Make a scatterplot: N300 & STAI-S
```{r}
ggplot(PM, aes(x = N300, y = STAIS)) +
  geom_point() +
    geom_smooth(method = "lm")
```

## Make a scatterplot: N300 & STAI-T
```{r}
ggplot(PM, aes(x = N300, y = STAIT)) +
  geom_point() +
    geom_smooth(method = "lm")
```

## Make a scatterplot: PP & BAI
```{r}
ggplot(PM, aes(x = PP, y = BAI)) +
  geom_point() +
    geom_smooth(method = "lm")
```

## Make a scatterplot: PP & STAI-S
```{r}
ggplot(PM, aes(x = PP, y = STAIS)) +
  geom_point() +
    geom_smooth(method = "lm", color = "dark green") +
  labs(title="Prospective Positivity and State Anxiety",
       x="Prospective Positivity Amplitude (mV)", y = "STAI-S Score")+
  theme_classic()  +
  theme(text = element_text(size = 20))                    
```

## Make a scatterplot: PP & STAI-T
```{r}
ggplot(PM, aes(x = PP, y = STAIT)) +
  geom_point() +
    geom_smooth(method = "lm")
```

## select variables of interest
```{r}
PM_N300 <- select(PM, N300, STAIS, STAIT, BAI)
PM_PP <- select(PM, PP, STAIS, STAIT, BAI)
```


## Get correlation of variables
### Via cor() function
```{r}
cor(PM_N300)
cor(PM_PP)

cor.test( formula = ~ N300 + STAIS,
          data = PM_N300)
cor.test( formula = ~ N300 + STAIT,
          data = PM_N300)
cor.test( formula = ~ N300 + BAI,
          data = PM_N300)

cor.test( formula = ~ PP + STAIS,
          data = PM_PP)
cor.test( formula = ~ PP + STAIT,
          data = PM_PP)
cor.test( formula = ~ PP + BAI,
          data = PM_PP)
```

### Via apa.cor.table()
```{r}
apa.cor.table(PM_N300, "PM_N300_anxiety.doc", show.conf.interval = TRUE)
apa.cor.table(PM_PP, "PM_PP_anxiety.doc", show.conf.interval = TRUE)
```

## FDR correction
```{r}
# Vector of unadjusted p values 

p2 <- c(0.6255, 0.7373, 0.4192, 0.0207, 0.141, 0.1307)

# Key:  N300 + STAIS, N300 + STAIT, N300 + BAI, PP + STAIS, PP + STAIT, PP + BAI

# FDR correction

FDR_Hyp2 <- p.adjust(p2, method = "fdr", n = length(p2))

FDR_Hyp2
```

# Hypothesis 3

## Make a scatterplot: N300 & Accuracy
```{r}
ggplot(PM, aes(x = N300, y = Accuracy)) +
  geom_point() +
    geom_smooth(method = "lm")
```
## Make a scatterplot: N300 & RT
```{r}
ggplot(PM, aes(x = N300, y = RT)) +
  geom_point() +
    geom_smooth(method = "lm")
```


## Make a scatterplot: PP & Accuracy
```{r}
ggplot(PM, aes(x = PP, y = Accuracy)) +
  geom_point() +
    geom_smooth(method = "lm", color = "dark green") +
 labs(title="Prospective Positivity and Accuracy",
       x="Prospective Positivity Amplitude (mV)", y = "Accuracy")+
  theme_classic() +
 theme(text = element_text(size = 20)) 
```

## Make a scatterplot: PP & RT
```{r}
ggplot(PM, aes(x = PP, y = RT)) +
  geom_point() +
    geom_smooth(method = "lm")
```


## select variables of interest
```{r}
PM_N300_acc_RT <- select(PM, N300, Accuracy, RT)
PM_PP_acc_RT <- select(PM, PP, Accuracy, RT)
```

## Get correlation of variables
### Via cor() function
```{r}
cor(PM_N300_acc_RT)
cor(PM_PP_acc_RT)

cor.test( formula = ~ N300 + Accuracy,
          data = PM_N300_acc_RT)
cor.test( formula = ~ N300 + RT,
          data = PM_N300_acc_RT)

cor.test( formula = ~ PP + Accuracy,
          data = PM_PP_acc_RT)
cor.test( formula = ~ PP + RT,
          data = PM_PP_acc_RT)
```

### Via apa.cor.table()
```{r}
apa.cor.table(PM_N300_acc_RT, "PM_N300_acc.doc", show.conf.interval = TRUE)
apa.cor.table(PM_PP_acc_RT, "PM_PP_RT.doc", show.conf.interval = TRUE)
```

## FDR correction
```{r}
# Vector of unadjusted p values 

p3 <- c(0.07467, 0.2091, 0.03479, 0.9157)

# Key: N300 + Accuracy, N300 + RT, PP + Accuracy, PP + RT

# FDR correction

FDR_Hyp3 <- p.adjust(p3, method = "fdr", n = length(p3))

FDR_Hyp3
```

# Mediation

## Load libraries
```{r}
#install.packages("lavaan", dependencies = TRUE)
#install.packages("lavaanPlot")
library(lavaan)
library(lavaanPlot)
```

## Consider a classical mediation setup with three variables: Y is the dependent variable, X is the predictor, and M is a mediator. Fit a path analysis model that includes the direct effect of X on Y and the indirect effect of X on Y via M
```{r}
# Y = Accuracy (accuracy on prospective memory task)
# X = STAIS (state anxiety)
# M = PP (prospective positivity amplitude)

PM_Mediation <- select(PM, Accuracy, STAIS, PP)

set.seed(1234)
mediation.model <- ' # direct effect
             Accuracy ~ c*STAIS
           # mediator
             PP ~ a*STAIS
             Accuracy ~ b*PP 
           # indirect effect (a*b)
             ab := a*b
           # total effect
             total := c + (a*b)
         '
fit <- sem(mediation.model, data = PM_Mediation, meanstructure = TRUE,
           se = "boot", bootstrap = 1000)
summary(fit, fit.measures=T, standardized=T, ci=TRUE)

parameterEstimates(fit, 
                   boot.ci.type = "bca.simple",
                   level = .95, ci = TRUE,
                   standardized = TRUE)

# SEM plot
labels <- list(Accuracy = "Prospective Memory Accuracy", STAIS = "State Anxiety", PP = "Prospective Positivity Amplitude")

lavaanPlot(model = fit, labels = labels, node_options = list(shape = "box", 
           fontname = "times"), edge_options = list(color = "grey"), 
           coefs = TRUE, covs = TRUE, stars = "regress")



# Estimate = raw regression coefficient (i.e., b)
# Std. Err = error for the estimate
# Z-value = Estimate / Std. Err
# P(>|z|) = significance value
# ci.lower/upper = 95% CI for the estimate
# Std.lv = …just ignore this one
# Std.all = standardized beta coefficient

```

# T Tests and Correlations on extra variables

#Split up groups based on anxiety screener
```{r}

HighA <- subset(PM, ID == 'PMA01'| ID == 'PMA02' | ID == 'PMA03'| ID == 'PMA06' | ID == 'PMA08' | ID == 'PMA11'| ID == 'PMA15' | ID == 'PMA16' | ID == 'PMA17' | ID == 'PMA18' | ID == 'PMA20'| ID == 'PMA22' | ID == 'PMA24' | ID == 'PMA25' | ID == 'PMA27' | ID == 'PMA28' | ID == 'PMA37' | ID == 'PMA42' | ID == 'PMA43' | ID == 'PMA44' | ID == 'PMA45'| ID == 'PMA47' | ID == 'PMA49'| ID == 'PMA52' | ID == 'PMA58' | ID == 'PMA61'| ID == 'PMA65' | ID == 'PMA66' | ID == 'PMA67' | ID == 'PMA68' | ID == 'PMA75')

LowA <- subset(PM, ID == 'PMA04'| ID == 'PMA07' | ID == 'PMA10'| ID == 'PMA12' | ID == 'PMA13' | ID == 'PMA32'| ID == 'PMA35' | ID == 'PMA36' | ID == 'PMA38' | ID == 'PMA41' | ID == 'PMA46' | ID == 'PMA48' | ID == 'PMA50' | ID == 'PMA51' | ID == 'PMA53' | ID == 'PMA54' | ID == 'PMA55' | ID == 'PMA56' | ID == 'PMA57' | ID == 'PMA59' | ID == 'PMA60' | ID == 'PMA62' | ID == 'PMA69' | ID == 'PMA70' | ID == 'PMA72' | ID == 'PMA73' | ID == 'PMA74' | ID == 'PMA76' | ID == 'PMA77' | ID == 'PMA78')

```


## Sleepiness
```{r}
library(lsr)

t_test_result_sleepiness = t.test(HighA$Sleepiness, LowA$Sleepiness, var.equal = T, paired=FALSE)

t_test_result_sleepiness

cohensD(HighA$Sleepiness, LowA$Sleepiness)

cor.test( formula = ~ Sleepiness + BAI,
          data = PM)

cor.test( formula = ~ Sleepiness + STAIS,
          data = PM)

cor.test( formula = ~ Sleepiness + STAIT,
          data = PM)

cor.test( formula = ~ Sleepiness + Accuracy,
          data = PM)

cor.test( formula = ~ Sleepiness + RT,
          data = PM)

cor.test( formula = ~ Sleepiness + N300,
          data = PM)

cor.test( formula = ~ Sleepiness + PP,
          data = PM)

```

## Hours slept
```{r}
t_test_result_SleepHours = t.test(HighA$SleepHours, LowA$SleepHours, var.equal = T, paired=FALSE)

t_test_result_SleepHours

cohensD(HighA$SleepHours, LowA$SleepHours)

cor.test( formula = ~ SleepHours + BAI,
          data = PM)

cor.test( formula = ~ SleepHours + STAIS,
          data = PM)

cor.test( formula = ~ SleepHours + STAIT,
          data = PM)

cor.test( formula = ~ SleepHours + Accuracy,
          data = PM)

cor.test( formula = ~ SleepHours + RT,
          data = PM)

cor.test( formula = ~ SleepHours + N300,
          data = PM)

cor.test( formula = ~ SleepHours + PP,
          data = PM)

```

## ADHD
```{r}
t_test_result_ADHD = t.test(HighA$ADHD, LowA$ADHD, var.equal = T, paired=FALSE)

t_test_result_ADHD

cor.test( formula = ~ ADHD + BAI,
          data = PM)

cor.test( formula = ~ ADHD + STAIS,
          data = PM)

cor.test( formula = ~ ADHD + STAIT,
          data = PM)

cor.test( formula = ~ ADHD + Accuracy,
          data = PM)

cor.test( formula = ~ ADHD + RT,
          data = PM)

cor.test( formula = ~ ADHD + N300,
          data = PM)

cor.test( formula = ~ ADHD + PP,
          data = PM)

```


## Depression
```{r}
t_test_result_Depression = t.test(HighA$Depression, LowA$Depression, var.equal = T, paired=FALSE)

t_test_result_Depression

cohensD(HighA$Depression, LowA$Depression)

cor.test( formula = ~ Depression + BAI,
          data = PM)

cor.test( formula = ~ Depression + STAIS,
          data = PM)

cor.test( formula = ~ Depression + STAIT,
          data = PM)

cor.test( formula = ~ Depression + Accuracy,
          data = PM)

cor.test( formula = ~ Depression + RT,
          data = PM)

cor.test( formula = ~ Depression + N300,
          data = PM)

cor.test( formula = ~ Depression + PP,
          data = PM)

```

## Positive Affect
```{r}
t_test_result_PositiveAff = t.test(HighA$PostiveAff, LowA$PostiveAff, var.equal = T, paired=FALSE)

t_test_result_PositiveAff

cohensD(HighA$PostiveAff, LowA$PostiveAff)

cor.test( formula = ~ PostiveAff + BAI,
          data = PM)

cor.test( formula = ~ PostiveAff + STAIS,
          data = PM)

cor.test( formula = ~ PostiveAff + STAIT,
          data = PM)

cor.test( formula = ~ PostiveAff + Accuracy,
          data = PM)

cor.test( formula = ~ PostiveAff + RT,
          data = PM)

cor.test( formula = ~ PostiveAff + N300,
          data = PM)

cor.test( formula = ~ PostiveAff + PP,
          data = PM)

```


## Negative Affect
```{r}
t_test_result_NegativeAff = t.test(HighA$NegativeAff, LowA$NegativeAff, var.equal = T, paired=FALSE)

t_test_result_NegativeAff

cohensD(HighA$NegativeAff, LowA$NegativeAff)

cor.test( formula = ~ NegativeAff + BAI,
          data = PM)

cor.test( formula = ~ NegativeAff + STAIS,
          data = PM)

cor.test( formula = ~ NegativeAff + STAIT,
          data = PM)

cor.test( formula = ~ NegativeAff + Accuracy,
          data = PM)

cor.test( formula = ~ NegativeAff + RT,
          data = PM)

cor.test( formula = ~ NegativeAff + N300,
          data = PM)

cor.test( formula = ~ NegativeAff + PP,
          data = PM)
```

## Motivation
```{r}
t_test_result_Motivation = t.test(HighA$Motivation, LowA$Motivation, var.equal = T, paired=FALSE)

t_test_result_Motivation

cohensD(HighA$Motivation, LowA$Motivation)

cor.test( formula = ~ Motivation + BAI,
          data = PM)

cor.test( formula = ~ Motivation + STAIS,
          data = PM)

cor.test( formula = ~ Motivation + STAIT,
          data = PM)

cor.test( formula = ~ Motivation + Accuracy,
          data = PM)

cor.test( formula = ~ Motivation + RT,
          data = PM)

cor.test( formula = ~ Motivation + N300,
          data = PM)

cor.test( formula = ~ Motivation + PP,
          data = PM)

```


## DEX
```{r}
cor.test( formula = ~ DEX + BAI,
          data = PM)

cor.test( formula = ~ DEX + STAIS,
          data = PM)

cor.test( formula = ~ DEX + STAIT,
          data = PM)

cor.test( formula = ~ DEX + Depression,
          data = PM)

cor.test( formula = ~ DEX + Depression,
          data = PM)

cor.test( formula = ~ DEX + Depression,
          data = PM)

cor.test( formula = ~ DEX + Accuracy,
          data = PM)

cor.test( formula = ~ DEX + RT,
          data = PM)

cor.test( formula = ~ DEX + N300,
          data = PM)

cor.test( formula = ~ DEX + PP,
          data = PM)


```

## Partial correlations using pcor.test() 

### Create formula to get confidence intervals for pcor.test()
```{r}
pcor_ci.test <-
function (x, y, z, method = c("pearson", "kendall", "spearman"), conf.level = 0.95, ...) {
    d1 <- deparse(substitute(x))
    d2 <- deparse(substitute(y))
    d3 <- deparse(substitute(z))
    data.name <- paste0(d1, " and ", d2, "; controlling: ", d3)
    method <- match.arg(method)
    Method <- paste0("Partial correlation (", method, ")")
    alternative <- "true partial correlation is not equal to 0"

    x <- as.vector(x)
    y <- as.vector(y)
    z <- as.data.frame(z)
    xyz <- data.frame(x, y, z)
    pcor <- ppcor::pcor(xyz, method = method)
    estimate <- pcor$est[1, 2]
    p.value <- pcor$p.value[1, 2]
    parameter <- c(n = pcor$n, gp = pcor$gp)
    statistic <- c(Stat = pcor$statistic[1, 2])

    fit1 <- lm(x ~ z, data = xyz)
    fit2 <- lm(y ~ z, data = xyz)
    cortest <- cor.test(resid(fit1), resid(fit2), method = method, conf.level = conf.level, ...)
    ci <- cortest$conf.int

    ht <- list(
        statistic = statistic,
        parameter = parameter,
        p.value = p.value,
        estimate = c(partial.cor = estimate),
        alternative = alternative,
        method = Method,
        data.name = data.name,
        conf.int = ci
    )
    class(ht) <- "htest"
    ht
}
```


### Anxiety & Prospective Positivity - negative affect
```{r}
pcor.test(PM$STAIS, PM$PP, PM$NegativeAff, method = c("pearson"))
pcor_ci.test(PM$STAIS, PM$PP, PM$NegativeAff,  method = c("pearson")) 

pcor.test(PM$STAIT, PM$PP, PM$NegativeAff, method = c("pearson"))
pcor_ci.test(PM$STAIT, PM$PP, PM$NegativeAff,  method = c("pearson")) 

pcor.test(PM$BAI, PM$PP, PM$NegativeAff, method = c("pearson"))
pcor_ci.test(PM$BAI, PM$PP, PM$NegativeAff,  method = c("pearson")) 
```


### Anxiety & Prospective Positivty - sleepiness
```{r}

pcor.test(PM$STAIS, PM$PP, PM$Sleepiness, method = c("pearson"))
pcor_ci.test(PM$STAIS, PM$PP, PM$Sleepiness,  method = c("pearson")) 

pcor.test(PM$STAIT, PM$PP, PM$Sleepiness, method = c("pearson"))
pcor_ci.test(PM$STAIT, PM$PP, PM$Sleepiness,  method = c("pearson")) 

pcor.test(PM$BAI, PM$PP, PM$Sleepiness, method = c("pearson"))
pcor_ci.test(PM$BAI, PM$PP, PM$Sleepiness,  method = c("pearson")) 

```



## Anxiety & N300 - sleepiness 
```{r}
pcor_ci.test(PM$STAIS, PM$N300, PM$Sleepiness,  method = c("pearson")) 
pcor.test(PM$STAIS, PM$N300, PM$Sleepiness,  method = c("pearson")) 

pcor_ci.test(PM$STAIT, PM$N300, PM$Sleepiness,  method = c("pearson")) 
pcor.test(PM$STAIT, PM$N300, PM$Sleepiness, method = c("pearson"))

pcor_ci.test(PM$BAI, PM$N300, PM$Sleepiness,  method = c("pearson")) 
pcor.test(PM$BAI, PM$N300, PM$Sleepiness, method = c("pearson"))

```


## Anxiety and accuracy - sleepiness 
```{r}
pcor_ci.test(PM$STAIS, PM$Accuracy, PM$Sleepiness, method = c("pearson"))
pcor.test(PM$STAIS, PM$Accuracy, PM$Sleepiness,  method = c("pearson")) 

pcor_ci.test(PM$STAIT, PM$Accuracy, PM$Sleepiness, method = c("pearson"))
pcor.test(PM$STAIT, PM$Accuracy, PM$Sleepiness,  method = c("pearson")) 

pcor_ci.test(PM$BAI, PM$Accuracy, PM$Sleepiness, method = c("pearson"))
pcor.test(PM$BAI, PM$Accuracy, PM$Sleepiness,  method = c("pearson")) 

```

## N300 and accuracy - sleepiness 
```{r}

pcor_ci.test(PM$N300, PM$Accuracy, PM$Sleepiness, method = c("pearson"))
pcor.test(PM$N300, PM$Accuracy, PM$Sleepiness,  method = c("pearson")) 

```

## PP and accuracy - sleepiness 
```{r}
pcor_ci.test(PM$PP, PM$Accuracy, PM$Sleepiness, method = c("pearson"))
pcor.test(PM$PP, PM$Accuracy, PM$Sleepiness,  method = c("pearson")) 
```


# Tables

### Via apa.cor.table()
```{r}
apa.cor.table(PM, "PM_all.doc", show.conf.interval = TRUE)

 PM_explore1 <- select(PM, BAI, STAIS, STAIT, SleepHours, Sleepiness, Depression, PostiveAff, NegativeAff, Motivation, ADHD)

 apa.cor.table(PM_explore1, "PM_explore1.doc", show.conf.interval = TRUE)

 PM_explore2 <- select(PM, Accuracy, RT, SleepHours, Sleepiness, Depression, PostiveAff, NegativeAff, Motivation, ADHD)

 apa.cor.table(PM_explore2, "PM_explore2.doc", show.conf.interval = TRUE)

PM_explore3 <- select(PM, N300, PP, SleepHours, Sleepiness, Depression, PostiveAff, NegativeAff, Motivation, ADHD)

 apa.cor.table(PM_explore3, "PM_explore3.doc", show.conf.interval = TRUE)

 PM_explore_combined <- select(PM, Accuracy, RT, N300, PP, SleepHours, Sleepiness, Depression, PostiveAff, NegativeAff, Motivation, ADHD)

 apa.cor.table(PM_explore_combined, "PM_explore_combined.doc", show.conf.interval = TRUE)

# apa.cor.table()
```







